# Intro

This is a short guide on how to use the new data quality matrix adding/upating
features of `PAMscapes`. This mostly consists of the new `addDataQuality` function
and the existing `runDailyLTSAReview` function. This requires `PAMscapes` version
0.17.0 or later, which can be installed from GitHub.

This file and the files used for the examples can be found on NOAA Google Cloud
Storage at "pam-ww-tmp/Taiki_Sakai/DQ"

```{r}
if(packageVersion('PAMscapes') < '0.17.0') {
    devtools::install_github('TaikiSan21/PAMscapes')
}
```

`addDataQuality` is a new tool that allows you to provide time-frequency annotations
to a set of NetCDF files to update the data quality matrix. If the NetCDF files
do not yet have a data quality matrix, it will create a new one. Annotations can
either come directly from the Makara "recording_intervals" table, or can be provided
as a dataframe with columns `start`, `end`, `freqMin`, `freqMax` and `quality` where
the time values must be in POSIXct format (or YYYY-MM-DD HH:MM:SS for conversion) and
in UTC time zone, and frequencies are in Hz. There can also be an optional column
used to match certain annotations using the "platform" global attribute of a NetCDF
file - this can be useful to be able to store annotations for multiple deployments
in a single file, even if deployments overlap in time. 

Annotations can also be created and saved using the `runDailyLTSAReview` Shiny app
tool, then passed in to `addDataQuality`. This tool allows user to page through
NetCDF files and draw boxes indicating areas of interest.

Start by loading some data that do not have updated quality data.

```{r}
library(PAMscapes)
ncFiles <- c('nc/NEFSC_GOM_202309_USTR01_20231028.nc', 
                  'nc/NEFSC_GOM_202309_USTR01_20231029.nc',
                  'nc/NEFSC_GOM_202309_USTR01_20231030.nc')
dataOriginal <- loadSoundscapeData(ncFiles)
```

From the LTSA plot we can see that there is a sudden change in the
data before 10-29 12:00 that seems like an instrument problem.

```{r}
plotLTSA(dataOriginal)
```
## Function Parameters

`addDataQuality` has the following parameters:

* `x` - a NetCDF, vector of files, or folder containing files. If a folder, then
all .nc files directly within that folder will be used (subfolders will not be searched)
* `annotation` - a dataframe listing time-frequency bounds. Time bounds must be
in POSIXct format (or YYYY-MM-DD HH:MM:SS characters) in UTC timezone in columns
`start` and `end`. Frequency bounds must be in Hz in columns `freqMin` and `freqMax`.
Numeric quality code should be in column `quality`. 
* `platformColumn` - NetCDF files should contain a "platform" global attribute,
this can be used to specify that certain rows of `annotation` only apply to certain
files in `x`. This value specifies the name of the column in `annotation` should 
be used (e.g. `platformColumn="deployment_code"` means `annotation$deployment_code` 
contains values that will match the "platform" global attribute)
* `updateUnknown` - The numeric quality code to update "Unknown (2)" quality codes
in the data quality matrix with. Most commonly set to `1`, or can be `FALSE` if you
do not want to change the `2` values
* `preview` - If this is set to `TRUE`, then no changes will be made to the files.
Instead only plots will be made so that you can verify that changes look appropriate
before making edits to the files. If `FALSE` then the files in `x` will be edited.
* `plot` - If `TRUE` then plots of the data quality matrix will be created for
all files after annotations have been applied. If `plot="new"` then plots will
only be created for files that contained an annotation, this can be better when
applying few annotations to a large number of files.
* `reset` - If `TRUE`, then all values of the data quality matrix will be 
reset to `2` before applying any annotations. This is primarily used in cases
where an incorrect annotation was previously applied

## Makara Annotation

Makara has a "recording_intervals" table that exists for the purpose of marking out
time-frequency ranges as either "COMPROMISED" or "UNUSABLE". Here's an example of some
output from this table.

```{r}
library(dplyr)
makaraIntervals <- read.csv('makara_intervals.csv', stringsAsFactors = FALSE) %>% 
    select(deployment_code,
           recording_interval_start_datetime,
           recording_interval_end_datetime,
           recording_interval_min_frequency_khz,
           recording_interval_max_frequency_khz,
           recording_interval_quality_type_code)
str(makaraIntervals)
```

`addDataQuality` is designed to work easily with this Makara output. It 
knows what the Makara column names mean, and that the frequency values are in
kHz instead of Hz. It also knows that the "deployment_code" column should be used
to match to the "platform" global attribute. 

In this case the deployment_code "NEFSC_GOM_202309_USTR01" as an annotation marking
the louder section we observed above as "UNUSABLE (4)", and the remaining data are set
to "GOOD (1)"

```{r}
qs <- addDataQuality(ncFiles, 
                     annotation = makaraIntervals, 
                     preview=TRUE,
                     updateUnknown = 1,
                     plot=TRUE)
```

## Dataframe Annotation

Annotations do not have to come from Makara, users can specify their own by
following some basic formatting. Columns `start` and `end` in the dataframe
should specify the time bounds, and must either be already in POSIXct format
or a character in YYYY-MM-DD HH:MM:SS, and timezone must be UTC. Columns
`freqMin` and `freqMax` should specify the frequency bounds, and are in units 
of Hz. Column `quality` should specify the quality code to assign.

If both frequencies are NA, all frequencies between times will be marked.
If both times are NA, all times between frequencies will be marked.

Here we can see that the "3" annotation has been applied to all frequencies
and the "4" annotation has been applied to all times, while the "2"
annotation has been applied to the time-frequency box.

```{r}
annoDf <- data.frame(
    start=c('2023-10-29 05:00:00', NA, '2023-10-30 05:00:00'),
    end = c('2023-10-30 01:00:00', NA, '2023-10-30 11:00:00'),
    freqMin = c(NA, 100, 1000),
    freqMax = c(NA, 500, 3000),
    quality=c(3, 4, 2)
)
addDataQuality(ncFiles, 
               annotation = annoDf, 
               preview=TRUE,
               updateUnknown = 1, 
               plot=TRUE)
```

Additionally, a separate column can be used to mark which rows should match
the "platform" global attribute. For this data, the "platform" is 
"NEFSC_GOM_202309_USTR01", so we can create a `deployment` column in the
annotation, and then add `platformColumn = "deployment"` to `addDataQuality`
so that only the first two annotations are applied to this data, and the
annotation for "SITE_B" is not.

```{r}
annoDf$deployment <- c('NEFSC_GOM_202309_USTR01', 'NEFSC_GOM_202309_USTR01', 'SITE_B')
addDataQuality(ncFiles, 
               annotation = annoDf,
               platformColumn = 'deployment',
               preview=TRUE,
               updateUnknown = 1, 
               plot=TRUE)
```

## Shiny App Annotation

The `runDailyLTSAReview` function opens up a Shiny app that lets you browse
through individual files and draw boxes around areas of interest, assigning
them a quality code. These can be saved by hitting the "Download Annotations"
button in the "Annotations" tab of the app.

```{r}
runDailyLTSAReview(ncFiles, plotQuality=TRUE)
```

The file from the app got saved as "DailyAnnotations.csv", and can be read 
directly into `addDataQuality`. Here we use the option `plot="new"` to only
show the quality matrix plot for the files that had annotations (only one day)
instead of all three files.

```{r}
appAnno <- read.csv('DailyAnnotations_Example.csv')
addDataQuality(ncFiles,
               annotation = appAnno,
               preview=TRUE,
               updateUnknown = 1,
               plot='new'
)
```

## Applying Annotations

All examples so far have used `preview=TRUE` so that changes were not applied to
the data, so lets apply the annotations from Makara - these should mark the loud 
sections of the soundscape data as "4 (Unusable)". When these data get read in
with `loadSoundscapeData`, any data points marked with quality flag of 3 or 4 
get set to NA values.

Plotting the LTSA of the updated data shows that region as gray (NA). The PSD
of the original data is completely dominated by the problematic region, and
the PSD of the updated data is much better.

The code at the bottom of this chunk shows how to reset the quality matrix data
in the files with `annotation=NULL`, `reset=TRUE`, and `preview=FALSE` so that
the examples in this document can be run again without the annotations applied.

```{r}
addDataQuality(ncFiles, 
               annotation=makaraIntervals,
               updateUnknown=1,
               preview=FALSE
)
dataUpdated <- loadSoundscapeData(ncFiles)
plotLTSA(dataOriginal, title='Original')
plotLTSA(dataUpdated, title='DQ Added, 4s Dropped by PAMscapes')
plotPSD(dataOriginal, title='Original')
plotPSD(dataUpdated, title='DQ Added, 4s Dropped by PAMscapes')
# reset files so examples can be run again
addDataQuality(ncFiles, 
               annotation=NULL,
               reset=TRUE,
               preview=FALSE
)
```
