# Intro

Demo using PAMscapes v 0.17.1 currently only on GitHub, available on CRAN
shortly after this demo (2026-09-24)

```{r}
devtools::install_github('TaikiSan21/PAMscapes')
```

The files for this demo are available on the NMFS cloud at:
pam-ww-tmp/Taiki_Sakai/PAMSI_202609_Demo

```{r}
library(bigrquery)
library(dplyr)
library(dbplyr)
library(PAMscapes)
library(lubridate)
```

# Detections from Makara (BigQuery)

Now that Makara is up and running, lets see how we can access our data! We'll
use `dbplyr` to connect to the database and download some data, more examples
of how this works can be found here:

https://dbplyr.tidyverse.org/articles/dbplyr.html

This uses `bigrquery` to interface with BigQuery, it should authenticate using
your NOAA e-mail. Cloud workstations default to using the workstation service
account for credentials, so when working on the cloud we first need to remove
these and re-authorize using our NOAA e-mail. I believe users need to request
to be added to the BigQuery approval list (https://makara-462670382708.us-east4.run.app/contact)

```{r}
# this first line is cloud only to avoid using workstation credentials
gargle::cred_funs_add(credentials_gce = NULL) 
bq_auth()
```

Connect to the database - you should see a "Connections" tab in RStudio
that lets you see the various tables

```{r}
con <- DBI::dbConnect(
    bigrquery::bigquery(),
    project='ggn-nmfs-pacm-dev-1',
    dataset='makara'
)
```

The tables starting with "v_" or "view_" are pre-joined tables that have already
connected the metadata pieces from other tables, most users will only interact
with these tables. If you need something not present in these views, either reach
out to Jeff or myself, or you can join data from the other tables on your own.
The "v_daily_detections" table has data summarised as daily presence, but there
is also a "v_detections" table with the original resolution data

```{r}
view_daily_detections <- tbl(con, 'v_detections_daily')
colnames(view_daily_detections)
```

# Makara and PAMscapes

`dbplyr` lets you interact with the tables as if they were a dataframe using
the normal `dplyr` functions, but it does not actually download the data as
a dataframe that you can use until you call `collect()`

```{r}
daily <- view_daily_detections %>% 
    filter(sound_source_code %in% c('RIWH', 'BLWH', 'SEWH', 'FIWH', 'HUWH'),
           year(date) >= 2014,
           analysis_organization_code == 'NEFSC',
    ) %>% 
    collect()
str(daily)
```

`loadDetectionData` is designed to work seamlessly with the Makara view tables
so we can load this raw data in directly

```{r}
dailyDets <- loadDetectionData(daily)
str(dailyDets)
```

From here the various summary plots are easy to make

```{r}
plotAcousticScene(dailyDets)
```

```{r}
plotDetectionBoxplot(dailyDets, 
                     group=c('species', 'deployment'),
                     facet='species',
                     bin='day/week',
                     combineYears=TRUE)
```

Disconnect from the database once we are done downloading data

```{r}
DBI::dbDisconnect(con)
```

# Environmental Data Integration

By default `PAMscapes` only loads positive detection data (it is easier for plotting
functions to assume all rows are positive detections), but we can tell it to load
other values with the `detectedValues` parameter, so we'll also load the "NOT_DETECTED"
times from the daily presence dataset. For this example we'll just use 2023 data
for a faster example.

```{r}
dets2023 <- loadDetectionData(daily, 
                              detectedValues = c('DETECTED', 'NOT_DETECTED')) %>% 
    filter(year(UTC) == 2023)
table(dets2023$detectedFlag)
```

The `matchEnvData` from `PAMmisc` lets use pull down environmental data from ERDDAP
servers. Here we load "analysed_sst" data from ERDDAP dataset "jplMURSST41", but this will work for
any ERDDAP dataset by supplying the dataset ID as `nc` and the desired variables
as `var` (or leave `var` out and all variables will be added.

`buffer` is used to create an average of the variable in a space-time window
in order of Longitude, Latitude, Time (secs). We use this to average the SST over
a day to match our daily presence resolution (`86400/2` before and after).

Data download takes time (~8 minutes), so skipping running code and just loading

```{r}
library(PAMmisc)
# dets2023 <- matchEnvData(dets2023, 
#                          nc = 'jplMURSST41', 
#                          var='analysed_sst', 
#                          buffer=c(0,0, 86400/2))
dets2023 <- readRDS('dets2023_sst.rds')
```

Keeping the "NOT_DETECTED" values lets us see if there might be a relationship
between present/absence times of animals and the daily average SST

```{r}
ggplot(dets2023) +
    geom_density(aes(x=analysed_sst_mean, col=detectedFlag)) +
    facet_wrap(~species)
```

# Soundscapes and Detections

Load soundscape data - commented out and reading in previously loaded data
to save time. HMD results are originally in 1-minute resolution, so we'll
set `timeBin='1day'` to convert them to daily median values to match
our daily presence time resolution

```{r}
ncDir <- 'Z:/DETECTOR_OUTPUT/PYTHON_SOUNDSCAPE_PYPAM/Raw/NEFSC_GOM/NEFSC_GOM_202309_USTR09/NC/'
# hmd <- loadSoundscapeData(ncDir, timeBin='1day')
hmd <- readRDS('GOM_202309_USTR09_DailyHMD.rds')
str(hmd)
```

Call indices are a common way to try and use soundscape metrics to try and 
find whales. The basic idea is to compare the ratio of a target level and a
background level, and if this value is greater than 1 then there may be an
animal calling.

We can use `createOctaveLevel` with `type='bb'` to create broadband sound level
measurements of these levels, and then create the index by taking their ratio.
By default, `createOctaveLevel` creates measures that add up the energy in the
designated frequency ranges, so we need to set `normalized=TRUE` so that each
output level is normalized by its frequency bandwidth (otherwise the larger
noise ranges would be much larger values just by virtue of summing more data
points).

```{r}
bb <- createOctaveLevel(hmd, 
                        type='bb', 
                        freqRange=list(c(20,21), 
                                       c(12,34),
                                       c(42,43),
                                       c(36, 50)
                        ),
                        normalized=TRUE)

bb$finIndex <- bb$`BB_20-21` / bb$`BB_12-34`
bb$blueIndex <- bb$`BB_42-43` / bb$`BB_36-50`
plot(bb$finIndex, type='l')
lines(bb$blueIndex, col='blue')
```

Next we can use `matchDetectionData` to add the daily presence data from Makara
to these broadband levels. We add each species separately as a new column to
the broadband level. The soundscape data has a column "platform" that contains
the same deployment code stored in the "deployment" column of our detection data,
so we can set `by=c('platform' = 'deployment')` to let the code know that only
detections with deployment matching the platform value should be used (alternatively
you could filter `dailyDets` to only include data from that deployment first).

```{r}
bb <- matchDetectionData(bb, 
                         filter(dailyDets,  species == 'FIWH'),
                         name='finPresence',
                         by=c('platform' = 'deployment'),
                         value=TRUE,
                         fillNA=FALSE)
bb <- matchDetectionData(bb, 
                         filter(dailyDets, species == 'BLWH'),
                         name = 'bluePresence', 
                         by=c('platform' = 'deployment'),
                         value=TRUE,
                         fillNA=FALSE)
table(bb$finPresence)
table(bb$bluePresence)
```

Compare index to confirmed presence.

```{r}
ggplot(bb) +
    geom_density(aes(x=finIndex, col=finPresence)) +
    geom_vline(xintercept=1)
ggplot(bb) +
    geom_density(aes(x=blueIndex, col=bluePresence)) +
    geom_vline(xintercept=1)
```
