# PAMscapes GCloud Demo Day

Hi! Note that this demo document was created for an early cloud demo, and many
functions may now be out of date. Additionally, the paths to data present
will not function outside of that NMFS PAM Cloud environment.

## Install Stuff

`PAMscapes` 0.11.3 should be available on CRAN, so start by downloading from there.
We'll also use the `readxl` package later, and `dplyr` for messing with data.

```{r}
if(!require('PAMscapes') &&
   packageVersion('PAMscapes') < '0.11.3') {
    install.packages('PAMscapes')
    library(PAMscapes)
}
if(!require('readxl')) {
    install.packages('readxl')
    library(readxl)
}
if(!require('dplyr')) {
    install.packages('dplyr')
    library(dplyr)
}
packageVersion('PAMscapes') # veryify you get 0.11.3 installed
```
## Where is data

The cloud instances should have the `pam-ww-tmp` bucket already set up as
the `X` drive, check to make sure that exists. We'll be using data in the
`Cloud_Demo_Resources` folder. We'll also use some data from the `nefsc-1`,
which exists under `C:/pamdata_mount`

```{r}
dir.exists('X:/Cloud_Demo_Resources')
dir.exists('C:/pamdata_mount/nefsc-1')
```

## Soundscape Example

We'll load hybrid millidecade that was processed with PyPam.

```{r}
pypamDir <- 'C:/pamdata_mount/nefsc-1/bottom_mounted/NEFSC_SBNMS/NEFSC_SBNMS_202303_SB03/NEFSC_SBNMS_202303_SB03_ST/PYPAM/NC'
pypamFiles <- list.files(pypamDir, pattern='nc$', full.names=TRUE)
length(pypamFiles)
```

This takes about 1s / file to load, so we'll just take a look at a few. There's a
warning about HMD_24000 - no worries, sometimes bands at Nyquist exist but
should generally be ignored for analysis.

```{r}
pyData <- loadSoundscapeData(pypamFiles[2:10])
```

Easiest way to explore data is with this function that will launch a Shiny app 
to look at plots. App also creates code snippets so that if you like a plot / settings
you can easily generate it outside of the app.

```{r}
runSoundscapeExplorer(pyData)
```

Another useful functionality is generating third octave level or broadband level 
measurements from the HMD. This handles properly aligning / splitting HMD bands
into the desired output band.

```{r}
tol <- createOctaveLevel(pyData, type='tol', freqRange=c(100, 24e3))
str(tol)
bb <- createOctaveLevel(pyData, type='bb', freqRange=c(100, 1000))
str(bb)
```


## Detection Example

This work is still under development, so more work will be done in the near future.
Currently designed to work well with Makara data, some things are slightly
fudged for the purpose of this example.

```{r}
detData <- loadDetectionData('X:/Cloud_Demo_Resources/PAMscapes/Makara_Example.csv', source='csv', 
                           columnMap=PAMscapes:::getColMaps('makara'), extraCols='deployment')
str(detData)
```

Lets look at what we have!

```{r}
plotAcousticScene(detData)
```

Those names aren't exactly the most informative - this is relatively raw output
from the database, in the future we can easily/automatically join other metadata.
For now we'll have to fudge the part by pretending some Excel files are a database.

```{r}
library(readxl)
joinMakaraData <- function(x, org, global) {
    glob_source <- read_excel(global, sheet='sound_sources')[
        c('code', 'name', 'current_sound_source_code')
    ]
    names(glob_source) <- c('species_code', 'species', 'current_species_code')
    if('species' %in% names(x)) {
        x <- joinMakaraSpecies(x, glob_source)
    }
    if('call' %in% names(x)) {
        glob_call <- read_excel(global, sheet='call_types')[
            c('code', 'name')
        ]
    }

    org_deploy <- read_excel(org, sheet='deployments')[
        c('deployment_code', 'site_code', 'project_code', 'deployment_latitude', 'deployment_longitude',
          'organization_code')
    ]
    names(org_deploy) <- c('deployment', 'site', 'project', 'depLatitude', 'depLongitude', 'organization')
    if('deployment' %in% names(x)) {
        x <- left_join(x, org_deploy, by='deployment')
        if('Latitude' %in% names(x)) {
            x$Latitude[is.na(x$Latitude)] <- x$depLatitude[is.na(x$Latitude)]
        } else {
            x$Latitude <- x$depLatitude
        }
        if('Longitude' %in% names(x)) {
            x$Longitude[is.na(x$Longitude)] <- x$depLongitude[is.na(x$Longitude)]
        } else {
            x$Longitude <- x$depLongitude
        }
        x$depLatitude <- NULL
        x$depLongitude <- NULL
    }

    org_ana <- read_excel(org, sheet='analyses', readxl::cell_cols('B:O'))[
        c('deployment_code', 'analysis_sound_source_codes', 'analysis_start_datetime', 'analysis_end_datetime')
    ]
    names(org_ana) <- c('deployment', 'species_code', 'effortStart', 'effortEnd')


    org_ana <- PAMscapes:::spreadEffort(org_ana, commas='species_code')
    org_ana <- joinMakaraSpecies(org_ana, glob_source)
    if('deployment' %in% names(x)) {
        x <- left_join(x, org_ana, by=c('deployment', 'species'))
    }
    x
}

joinMakaraSpecies <- function(x, glob_source) {
    if('species' %in% names(x)) {
        x$species_code <- x$species
        x$species <- NULL
    }
    x <- left_join(x, glob_source, by='species_code')
    hasUpdate <- !is.na(x$current_species_code)
    if(any(hasUpdate)) {
        x$species_code[hasUpdate] <- x$current_species_code[hasUpdate]
        x$species <- NULL
        x$current_species_code <- NULL
        x <- left_join(x, glob_source, by='species_code')
    }
    x$current_species_code <- NULL
    noMatch <- is.na(x$species)
    if(any(noMatch)) {
        x$species[noMatch] <- x$species_code[noMatch]
    }
    x$species_code <- NULL
    x
}
```

These Excel files are basically the same data from various tables in the database,
we can add those in the future (note this takes a bit of time to run on the server)

```{r}
orgFile <- 'X:/Cloud_Demo_Resources/PAMscapes/Makara_Organization.xls'
globalFile <- 'X:/Cloud_Demo_Resources/PAMscapes/Makara_Global_References.xlsx'
detData <- joinMakaraData(detData, org=orgFile, glob=globalFile)
str(detData)
```

And now we have better names, as well as additional metadata like "who owns this data".

```{r}
plotAcousticScene(detData, by='organization')
```

Looking at just the SWFSC drifter data by location.

```{r}
detData %>% 
    filter(organization == 'SWFSC') %>% 
    plotAcousticScene(by='site')
```
And for SEFSC

```{r}
detData %>% 
    filter(organization == 'SEFSC') %>% 
    plotAcousticScene(by='site')
```

One obvious question from these would be are those gaps from lack of detection or lack
of effort. We can find out! We pulled some effort-related metadata from the database
earlier, so we just need to collect that and the function knows how to deal with it.
Note that we may even have different effort for different species.

```{r}
effort <- distinct(detData[c('effortStart', 'effortEnd', 'species', 'site', 'project', 'organization')])
names(effort) <- c('start', 'end', 'species', 'site', 'project', 'organization')

detData %>% 
    filter(organization == 'SEFSC') %>% 
    plotAcousticScene(by='site', effort=effort)
```

We can also show this same plot as ranges of frequencies for each species instead
of just a fixed bar (note the frequencies are made up, I'm a computer nerd not
a whale person).

```{r}
freqMap <- data.frame(
    type=c("Unid Pygmy/Dwarf Sperm Whale", "Blainville's Beaked Whale", "Gervais' Beaked Whale", "Goose-Beaked Whale", "Risso's Dolphin", "Unid Dolphin", "Sperm Whale"),
    freqMin=c(80e3, 50e3, 45e3, 30e3, 20e3, 15e3, 2e3),
    freqMax=c(120e3, 60e3, 55e3, 40e3, 30e3, 20e3, 10e3))
detData %>% 
    filter(organization == 'SEFSC') %>% 
    plotAcousticScene(by='site', effort=effort, freqMap=freqMap, freqMin=1e3)
```

There's also a boxplot for showing aggregated detections. Here we show the number of
days per week with detections, each different point making up the boxplot is a 
different site (click the button in top right to view larger)

```{r}
detData %>% 
    filter(organization == 'SEFSC') %>% 
    plotDetectionBoxplot(group=c('site'), facet='species')
```

That version doesn't have any "zero" detections because without effort data
we can't know if its no detections or no effort. We can add that.

```{r}
detData %>% 
    filter(organization == 'SEFSC') %>% 
    plotDetectionBoxplot(group=c('site'), facet='species', effort=effort)
```

There's also an option to combine all years of data into the same "year", so now each point
of the boxplot is a different site/year combination.

```{r}
detData %>% 
    filter(organization == 'SEFSC') %>% 
    plotDetectionBoxplot(group=c('site'), facet='species', combineYears=TRUE, effort=effort)
```

And you can change the time resolution you want to look at (this one takes some
processing time)

```{r}
detData %>% 
    filter(organization == 'SEFSC') %>% 
    plotDetectionBoxplot(group=c('site'), bin='hour/day', facet='species', combineYears=TRUE, effort=effort)

```

Its not just Makara data - we can handle data of any* format, just with a bit more
work.

```{r}
otherDetFile <- 'X:/Cloud_Demo_Resources/PAMscapes/Other_Detections.csv'
otherCsv <- read.csv(otherDetFile, stringsAsFactors = FALSE)
str(otherCsv)
```

There will be a Shiny app to help interactively sort out the relevant settings
to make this work for whatever data you have.

```{r}
otherDets <- loadDetectionData(otherDetFile,
                  source='csv', 
                  columnMap=list('UTC'='StartDate', 'Latitude'='LATITUDE_DDG_DEPLOYMENT',
                                 'Longitude'='LONGITUDE_DDG_DEPLOYMENT'),
                  detectionType='presence',
                  presenceDuration='1day',
                  wide=TRUE,
                  speciesCols=c('NARW_Occur', 'Fin_Occur', 'Humpback_Occur'),
                  detectedValues=c(1),
                  extraCols=c('SITE_NAME', 'PROJECT_NAME', 'PROJECT_DESCRIPTION')
)
str(otherDets)
plotDetectionBoxplot(otherDets, group='SITE_NAME', facet='species', combineYears=TRUE)
```

