---
title: "PAMscapes Tutorial"
author: "Taiki Sakai"
# output: html_document
output: pdf_document
date: 2023-08-04
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document introduces the plotting and summary functions included in the early
versions of the `PAMscapes` package (v0.7.0 as of this writing). This will not
cover a lot of the AIS download functions in depth since they are covered in 
a separate document `AISWorkflow.md` that should be available in the same location
as this file.

### Prerequisites

Currently (2024-10-02) `PAMscapes` v0.7.0 is available on CRAN. This tutorial also requires
`PAMpal` version 0.20.1 or greater and `PAMmisc` version 1.11.5 or greater, both of
which are also available on CRAN.

```{r, eval=FALSE}
install.packages('PAMscapes')
install.packages('PAMpal')
install.packages('PAMmisc')
```

This tutorial also requires AIS data in a folder `AIS_West`. You can download the
appropriate data by running through the `AISWorkflow.rmd` tutorial, available
in the same place as this document.

Load packages for the start of the examples:

```{r, message=FALSE}
library(here)
library(PAMscapes)
library(readr)
```

### Load and Format Data

For this example we will use drifter track data from one of the SWFSC ADRIFT buoy deployments. We also
have soundscape metrics calculated using Triton software, and we've already downloaded AIS data
following the guide in `AISWorkflow.rmd`. We'll use `readr::read_csv` to read in the various soundscape
and GPS data to handle type conversions for us, and use `readLocalAIS` to read in the AIS data we've
already downloaded. Note that the soundscape metrics provided here (`ADRIFT_017_OL_sim.csv` and 
`ADRIFT_017_BB_sim.csv`) are not the actual soundscape metrics, but synthetic data created to be similar
in appearance to the real soundscape metrics for demonstration purposes only. 

```{r}
gps <- read_csv(here('tutorial/ADRIFT_017_GPS.csv'), show_col_types = FALSE)
ol <- read_csv(here('tutorial/ADRIFT_017_OL_sim.csv'), show_col_types = FALSE)
bb <- read_csv(here('tutorial/ADRIFT_017_BB_sim.csv'), show_col_types = FALSE)
aisDir <- here('AIS_West')
ais <- readLocalAIS(gps, aisDir, distance=10e3)
```

Note that while this example
is using soundscape metrics in the format created by Triton, the functions shown here are general enough
to be used for metrics calculated using other methods. The formatting requirements for input data are
that the first column of the dataframe is `UTC`, formatted as `POSIXct` class with timezone `UTC`, and the
remainder of the columns storing the metrics are named using the convention `NAME_FREQUENCY`, where the
`NAME` is the same across all columns and the `FREQUENCY` is in Hertz (Hz). For example, data with columns
`UTC`, `OctLev_150`, `OctLev_200`, `OctLev_300` would be accepted.

Data created with [MANTA](https://doi.org/10.3389/fmars.2021.703650) and [PyPam](https://github.com/lifewatch/pypam) 
are also accepted even though they do not follow this convention. Typical MANTA CSV output has the first column 
as the date, then the
remaining columns named by just frequency, with no `NAME_` prefix. MANTA data processed with `PAMscapes`
will have the prefix `HMD_` ([hybrid millidecade](https://doi.org/10.1121/10.0003324)) 
appended to the frequency column names. NetCDF files from MANTA and PyPam are also accepted by `PAMscapes`
functions, users can load these as dataframes using the `loadSoundscapeData` function.

For NetCDF that contain metadata fields about the location and platform name of the instrument,
these fields will be read in as columns `Longitude`, `Latitude`, and `platform`.

```{r}
# Can load PyPam NetCDF using this
pypam <- loadSoundscapeData(here('tutorial/PyPamExample.nc'))
ol <- loadSoundscapeData(here('tutorial/ADRIFT_017_OL_sim.csv'))
cat(str(pypam, list.len=10))
str(ol)
tail(colnames(pypam))
```

#### Creating Third Octave Level & Octave Level Measurements

The soundscape metrics you have may be at a finer frequency resolution than is necessary for 
the analysis task at hand. Power Spectral Density (PSD) at a resolution of 1Hz or 
Hybrid Millidecade (HMD) outputs are common formats, but have much more frequency resolution
than most users needed. This means larger files, which can make working with them slower. 
Often it is enough to use the Third Octave Level (TOL) or Octave Level (OL) measurements to
observe patterns in the soundscape. `PAMscapes` includes a function `createOctaveLevel` that
allows users to to generate these octave level measurements from their finer scale inputs.

Here we'll show how this works for the small set of PyPam data we just loaded, which outputs in HMD
resolution.

```{r}
ncol(pypam)
```
The `createOctaveLevel` function takes this larger file as input, then we use the `type`
argument to specify OL (`type='ol'`) or TOL (`type='tol'`) bands.

```{r}
pypamOl <- createOctaveLevel(pypam, type='ol')
colnames(pypamOl)
```
This function will use the standard center frequencies for the OL/TOL bands, and by default
it will use all frequencies available in the input data. Octave bands are created by summing
(in linear space) the values of the input data that are within each octave band. This can
be changed to average the values in each band by setting `method='mean'`, but this is not
recommended.

In this data our frequencies start
at 10Hz, but we may not be interested in the full range of available frequencies. We can use
the `freqRange` argument to specify the upper and lower frequencies we want. Here we can
restrict our data to bands from 250Hz to 2000Hz

```{r}
pypamOl <- createOctaveLevel(pypam, type='ol', freqRange=c(250, 2000))
str(pypamOl)
```

Note that the `freqRange` parameter only checks against the *center* frequency of the
band, so users do not need to worry about making sure their specifications account for
all frequencies within the band, only that the desired center frequency is included.
Practically this means that `freqRange=c(200, 2000)` and `freqRange=c(250, 2000)`
are identical when calculating full octave bands since the lowest center frequency
in both is 250Hz.

This function can be used to generate octave level measures from any finer
resolution input - this means that we can even use TOL measures to calculate
OL measures, and expect nearly identical results (there may occasional be
band-edge differences but they should be minor).

```{r}
pypamTol <- createOctaveLevel(pypam, type='tol')
olFromTol <- createOctaveLevel(pypamTol, type='ol')
sum(abs(pypamOl[['OL_250']] - olFromTol[['OL_250']]))
```

Note that the standard hybrid-millidecade bands output by MANTA and PyPam are normalized
by their bandwidth. `createOctaveLevel` corrects for this properly before summing to
create the desired output bands, but by default the generated band level measurements
are *not* normalized by their bandwidth. This can be changed with the `normalized=TRUE`
option.

The plot shows that the un-normalized (blue) line is much higher than the original
hybrid millidecade output (black), which is expected because it has summed multiple
HMD values. The normalized ouput (red) closely tracks the original HMD output (it
does not track exactly because the OL_250 contains frequencies from 177 to 354, but
we are only plotting HMD_250).

```{r}
pypamOlNorm <- createOctaveLevel(pypam, type='ol', normalized=TRUE)
plot(x=pypam$UTC, y=pypam$HMD_250, type='l', col='black', ylim=c(70, 104))
lines(x=pypamOl$UTC, y=pypamOl$OL_250, col='blue')
lines(x=pypamOlNorm$UTC, y=pypamOlNorm$OL_250, col='red')
```


### Pairing Environmental and AIS Summary Data

With data in hand, lets use some of the `PAMscapes` functions to augment our GPS data with
some additional information. First we'll use `matchGFS` to add wind and precipitation rate
data to our GPS coordinates from the Global Forecast System (GFS) weather model. 
This will download relevant data from the 
[UCAR data server](https://rda.ucar.edu/datasets/ds084.1/) and pair it with the appropriate
GPS coordinates. The particular GFS datasets downloaded are the closest "forecast" datasets
to each point in time (e.g. \*.f000 or \*.f003). Other environmental data products from either ERDDAP or HYCOM data servers
can be added using the `matchEnvData` function from the `PAMmisc` package, more information
is available on the [PAMpal tutorial website](https://taikisan21.github.io/PAMpal/NextStepsProcessing.html#adding-environmental-data).

```{r, eval=FALSE}
gps <- matchGFS(gps)
```

This function has added columns:

- `windU` eastward wind velocity (meters / second)
- `windV` northward wind velocity (meters / second)
- `windMag` the total wind magnitude (meters / second)
- `precRate` average surface precipitation rate (kg / m^2 / s)
- `matchLong/Lat/Time` the coordinate value in the NetCDF file your data was matched to

```{r}
str(gps)
```

Now we'll use the `addAISSummary` function to add columns that summarise the AIS activity at each time in our GPS
track. This function is a little different than the `addAIS` function explained in detail in the other tutorial,
instead of adding all the raw AIS data (and creating many additional rows of data), this function just adds
summary information relating to how many ships are nearby at each point in time, and their speed and distance
if any are present. You can either supply AIS data that has already been read in with `readLocalAIS` to the `ais`
argument, or you can give it the directory containing your AIS data. The distance is used to decide whether or not
a ship is "nearby" your data for the purpose of counting nearby vessels, and should be the same as the `distance`
argument used in `readLocalAIS`.

This function adds columns:

- `nShips` the number of vessels within `distance` at this time
- `meanDist` the average distance (meters) to nearby ships (`NA` if `nShips` is 0)
- `meanSOG` the average speed (knots) over ground of the nearby ships (`NA` if `nShips` is 0)
- `closeDist` the distance (meters) to the closest nearby ship (`NA` if `nShips` is 0)
- `closeSOG` the speed over ground (knots) of the closest nearby ship (`NA` if `nShips` is 0)

```{r}
gps <- addAISSummary(gps, ais=ais, distance=10e3)
# This would give identical results
# gps <- addAISSummary(gps, ais=aisDir, distance=10e3)
str(gps)
```



### Plotting Functions

We'll use this augmented GPS data and the soundscape metric data we loaded earlier to explore the 
plotting functions available in `PAMscapes`.

#### plotTimeseries Function - Simple Timeseries

Let's start by examining some of our soundscape data. The `plotTimeseries` function 
lets you plot a simple timeseries of data, with a few extra possibilities. The input
is any data with a `UTC` column (it also accepts the `yyyy-mm-ddTHH:MM:SSZ` column
name from Triton). Then you specify the column (`column`) you want to plot, and a time bin (`bin`)
for displaying the data. Soundscape metrics are often calculated at a small timescale,
so binning the data to a larger timescale can be more useful to examine trends. This also
allows us to display confidence interval of measurements within that bin with the `q` 
argument. 


```{r}
plotTimeseries(bb, column='BB_100-24000', bin='1hour')
```

All plotting outputs in `PAMscapes` are `ggplot` objects, so they can be
manipulated after creation, and we can use the `patchwork` package to create 
multipanel plots. Here's a grid showing some of the various options of the plotting function.

```{r}
library(patchwork)
(plotTimeseries(bb, column='BB_100-24000', bin='1hour', q=.05) + 
     plotTimeseries(bb, column='BB_100-24000', bin='2hour', q=c(.1, .99))) /
    (plotTimeseries(bb, column='BB_100-24000', bin='1hour', units='Broadband dB Level') +
         plotTimeseries(bb, column='BB_100-24000', bin='1hour', title='Broadband Level Plot'))
```

#### plotTimeseries Heatmaps - Display Timeseries as Grid

`plotTimeseries` can also plot your timeseries data as a grid showing a heatmap of values instead
of just a simple line with `style='heatmap'`. 

```{r}
plotTimeseries(bb, column='BB_100-24000', bin='1hour', style='heatmap')
```

#### markNA Function - Removing Unwanted Data

The high values at the end of the broadband data are throwing off the color scales. There can be times
where there are known high-noise events that we want to exclude from our analysis, so `PAMscapes` has a
`markNA` function that lets us label certain time periods and frequency ranges as `NA` values so that
they will not display in plots. We just need to create a dataframe that has `start` and `end` times to
mark as `NA`.

```{r}
tail(bb, 20)
```

We can see that the spike occurs just after 12:18:00, so we'll use that for our time boundary.

```{r}
naDf <- data.frame(start=as.POSIXct('2022-04-28 12:19:00', tz='UTC'),
                   end = as.POSIXct('2022-04-28 12:43:00', tz='UTC'))
bb <- markNA(bb, na=naDf)
tail(bb, 20)
```

We can also change the timezone used for plots
using the `toTz` argument. Inputs are always expected to be UTC, but it can be useful to display
in local time.

```{r}
plotTimeseries(bb, column='BB_100-24000', bin='1hour', style='heatmap')/
    plotTimeseries(bb, column='BB_100-24000', bin='1hour', style='heatmap', toTz='America/Los_Angeles')
```

#### plotHourlyLevel Function - Time/Frequency Relationships

Another useful summary plot for soundscape metrics explores the distribution of sound levels
across time and frequency. `plotHourlyLevel` creates a plot with frequency on one axis, and
hour of the day on the other. Each point is the median level across the dataset. This can be
a great way to look for diurnal patterns in data across the range of frequencies.

```{r}
plotHourlyLevel(ol)
```

Frequency can also be displayed in linear scale by setting `scale='linear'`, and a different
minimum frequency for the plot can be set with the `freqMin` parameter. Like `plotTimeseries`
above there are also options for `title`, `units`, and `toTz` that can be set. Additionally
you can change the color palette used with the `cmap` option. 

```{r}
(plotHourlyLevel(ol, scale='log', freqMin=10) + 
     plotHourlyLevel(ol, scale='linear', cmap=scales::viridis_pal(option='B')(25))) /
    (plotHourlyLevel(ol, toTz='America/Los_Angeles') + 
         plotHourlyLevel(ol, title='Octave Level', units='Octave Level dB'))
```

#### plotPSD and prepPSDData - Plot Power Spectral Density Summary Across Frequency

One common way of summarising soundscape metrics is just to simply plot the distribution
of sound levels across frequency. `plotPSD` can plot this information either as just the
median level with optional quantiles (`style='quantile'`, similar to `plotTimeseries` above), or the full 
density distribution (`style='density'`).

```{r}
(plotPSD(pypam, style='quantile', q=.05)) /
    (plotPSD(pypam, style='density'))
```

PSD data can be quite large depending on the frequency and time resolution, so `PAMscapes`
provides the `prepPSDData` function so that users can load and process their data for plotting
(which can be time consuming for large files) separately from plotting. This means that it is
much faster to fiddle with plot settings since users aren't waiting for the data loading each
time they change the quantile level or color scheme.

```{r}
psdData <- prepPSDData(pypam)
(plotPSD(psdData, style='quantile', q=.1, color='orange') + 
        plotPSD(psdData, style='quantile', q=.2, color='red')) /
    (plotPSD(psdData, style='density', freqRange=c(100, 4e3)) +
         plotPSD(psdData, style='density', freqRange=c(100,1e3)))
```

Another important function of `prepPSDData` is that we can give it multiple files or dataframes
of data, and it will combine and summarise them. We may want to combine multiple large PSD files
from the same region into one plot, but dealing with many large files in `R` can be tricky. With
`prepPSDData` you can just give it a vector of file names and it will combine them for you so
that they can then be plotted. Alternatively you can give it a list of dataframes if you've
already loaded multiple files into `R` that you want to combine and plot. The only requirement
is that the data must all have measurements at the exact same frequency levels. Here we can try
to combine the two different octave level measurements we've made/loaded from this tutorial.

```{r}
combinedOl <- prepPSDData(list(ol, pypamOl))
```

We get a warning that they cannot be combined - even though the octave level measurements are present
at the same intervals, the `pypamOl` measurements stop at 4kHz. We can get around this by limiting `ol`
to the columns up to 4kHz.

```{r}
combinedOl <- prepPSDData(list(ol[1:6], pypamOl))
(plotPSD(ol, style='density', dbRange=c(50, 130)) +
    plotPSD(pypamOl, style='density', dbRange=c(50, 130))) /
    plotPSD(combinedOl, style='density', dbRange=c(50, 130))
```

This example is a little forced for the sake of the tutorial since it is inconvenient to
upload full resolution PSD data, and the plots look a little strange with octave level 
resolution data. Below is a sample code snippet of how you might use this functionality
for a folder of large PSD files.

```{r, eval=FALSE}
# sample workflow for use on a folder of PSD files - code is not meant to be run for tutorial
psdFolder <- 'folder/of/large/psd'
psdFiles <- list.files(psdFolder, pattern='csv$', full.names=TRUE)
# this can take a long time for large files
psdData <- prepPSDData(psdFiles)
# save this so you don't have to re-process in the future
saveRDS(psdData, 'CombinedPSDPlottingData.rds')
plotPSD(psdData)
```

There are a number of styling options for `plotPSD`. Like other plotting functions, we can set 
the range of frequencies shown with the `freqRange` parameter and the range of values shown 
with the `dbRange` parameter. We can also change the color of the line and quantile ribbon with 
`color` and the palette used for density with `cmap`. The `q` parameter for quantile plots
controls the ribbon shown. A value of 0 or 1 will show no ribbon, only the median line. A
length one `q` will plot a ribbon from `q` to `1-q`, e.g. `q=.05` will plot the 5th and
95th percent quantiles. A length 2 `q` will plot a ribbon between the specified bounds.

```{r}
(plotPSD(pypam, style='quantile', q=0, freqRange=c(100,1e4)) + 
     plotPSD(pypam, style='quantile', q=c(.01, .8), color='red', freqRange=c(100, 1e4))) /
    (plotPSD(pypam, style='density', cmap=scales::viridis_pal(option='B')(25), dbRange=c(30, 120)) +
         plotPSD(pypam, style='density', cmap=scales::viridis_pal(option='mako')(25), freqRange=c(1e2, 1e3)))
```

Labels for the Y-axis can also be changed with `units` and a title can be added with `title`, and the
frequency scale can be set to either linear or log (default).

```{r}
plotPSD(pypam, style='density', units='TOL dB re 1uPa', title='Linear Plot', scale='linear')
```

You can also plot the quantile lines overlaid on top of the density heatmap.

```{r}
plotPSD(pypam, style=c('quantile', 'density'), q=.05, color='red')
```

`plotPSD` also has the option to plot separate quantiles on the same plot using the `by` 
option. `by` specifies the column name that should be used for differentiating between 
the different types to plot. Note that this option only works if all data are already 
combined into a single dataframe, and also note that typically the presence of a extra 
column for labeling in the soundscape data will cause issues for other functions that
are not expecting it.

Here we'll some TOL data from PyPam on the same plot as some TOL metrics from Triton.
We need to specify the colors to use for the separate lines with the `color` argument,
either as a vector of the number of colors or as a palette function. We can also show
the quantile ribbons like normal.

```{r}
tolPypam <- createOctaveLevel(pypam, type='tol', freqRange=c(125, 4e3))
# we'll use the "Source" column for our labeling
tolPypam$Source <- 'PyPam'
tol <- loadSoundscapeData(here('tutorial/ADRIFT_017_TOL_sim.csv'))
tol$Source <- 'Triton'
# columns 15-24 contain frequency higher than the 2kHz upper limit of our
# PyPam data, so we'll remove them before combining
combined <- rbind(tolPypam, tol[-(15:24)])
plotPSD(combined, by='Source', color=c('blue', 'darkgreen')) /
    plotPSD(combined, by='Source', color=scales::hue_pal(), q=.05)
```

#### plotScaledTimeseries & PAMpal::timeJoin - Data With Different Ranges

Before using `plotScaledTimeseries`, we'll connect some of our soundscape metrics directly
to the GPS data we were working with earlier. This is easily done with the `timeJoin` function 
from the `PAMpal` package (v0.20.1 or later). The
Triton outputs we loaded earlier just need to have their time column changed to `UTC` so that we 
can use `PAMpal::timeJoin` to connect them. We'll only connect metrics that are within 1 hour of our
GPS points (`thresh=3600`). We'll also mark the same time range as `NA` in our octave level data
as we did to our broadband data earlier for consistency.

```{r, message=FALSE}
# requires v 0.20.1 or later
library(PAMpal)
colnames(bb)[1] <- 'UTC'
colnames(ol)[1] <- 'UTC'
ol <- markNA(ol, na=naDf)
gps <- timeJoin(gps, bb, thresh=3600)
gps <- timeJoin(gps, ol, thresh=3600)
str(gps)
```

Now we have soundscape metrics, AIS summaries, and environmental data all in the same dataframe.
It would be nice to plot these all together, but they have vastly different ranges of values. 
`plotScaledTimeseries` lets us plot different types of data on the same line plot, and will
rescale the data so that they are all visible on the same scale. We just tell it which
columns we want to look at (`columns`). This values of columns `BB_100-24000`, `nShips`, and
`windMag` are rescaled so that their minimum and maximum values match the min and max of the
first column, `OL_250`. This function is useful to see if there is a correlation between
the various types of data - e.g. does the noise level tend to increase when there are more
ships nearby.

```{r}
plotScaledTimeseries(gps, columns=c('OL_250', 'BB_100-24000', 'nShips', 'windMag'))
```

The plot above can be a bit cluttered and hard to read, so there are some options to help.
First, you can use the `lwd` parameter to set different linewidths for the different
columns. This can help to highlight the most important variable we want to compare things to.
This is a vector of length equal to the number of columns, specifying their widths in order
(note that the default `ggplot` linewidth is 0.5). Another option to reduce visual clutter
is to use the `relMax` option to make the rescaled columns only use a certain percentage 
of the available plot range. In the plot below `relMax=0.7` means that the 3 rescaled columns
only have maximum values that reach 70% of the available plot maximum.

```{r}
plotScaledTimeseries(gps, title='Linewidth', columns=c('OL_250', 'BB_100-24000', 'nShips', 'windMag'),
                     lwd=c(1, .5, .5, .5)) /
    plotScaledTimeseries(gps, title='RelMax', columns=c('OL_250', 'BB_100-24000', 'nShips', 'windMag'),
                     relMax=0.7)
```

Another potential downside to the rescaling is that the minimum value present in your data
may not be the most sensible lower range to show. In our plot above, it appears that `windMag`
starts at 0, then goes up, but really it just starts at whatever its minimum value is. The
`minVals` argument lets you specify a minimum value to use for plot scaling instead of
the minimum value present in your data. To increase visibility, we'll only plot two columns
here to showcase the difference. `minVals` is a vector of length equal to `columns`, and
a value of `NA` will use the minimum value present for that variable (this is the default
behavior). Here we plot one default (top), and one where we set the `windMag` lower bound
to 0. 

```{r}
plotScaledTimeseries(gps, columns=c('OL_250', 'windMag'), title='Bottom is Minimum') /
    plotScaledTimeseries(gps, columns=c('OL_250', 'windMag'), title='Bottom is Zero',
                         minVals=c(NA, 0))
```

The default colors for the different columns uses `scales::hue_pal` (which is the same
default color palette as `ggplot`, but this can be changed with the `color` argument.
You can either provide a different color palette generating function, or a vector
of color names equal to the number of `columns`.

```{r}
plotScaledTimeseries(gps, columns=c('OL_250', 'BB_100-24000', 'nShips', 'windMag'),
                     color=scales::viridis_pal()) /
    plotScaledTimeseries(gps, columns=c('OL_250', 'BB_100-24000', 'nShips', 'windMag'),
                     color=c('red', 'green', 'darkgreen', 'blue'))
```

#### plotAcousticScene - Frequency Contributions from Different Detections

The final type of plot displays the contributions of different detections in a
dataset to the total acoustic scene. It needs a dataframe that has some kind
of timestamped detections that are labeled as different types, and then
another dataframe specifying the frequency ranges that those detection
types correspond to. For this example we'll read in some `PAMpal` data to
fake an example plot. These echolocation click detections are classified to
a different click detector type with the `detectorName` column, so we'll use
that to show how the plotting function works. To reduce the size of the example
data, only the `UTC` and `detectorName` columns are included from a dataframe
that was created using the function `PAMpal::getClickData`. 

```{r}
clicks <- readRDS(here('tutorial/ADRIFT017_clicks.rds'))
table(clicks$detectorName)
```

First we'll need to create a dataframe that tells the function what frequency ranges to
associate with each value of `detectorName`. This needs to have column `type` that matches
the names of the types in the data we will plot, then `freqMin` and `freqMax` are the
frequency ranges for that type. We use this dataframe as the `freqMap` input for the
`plotAcousticScene` function. We also need to specify that the column within our data
that contains the different detection types we are interested in (`typeCol='detectorName'`).
The `bin` argument is the unit of time to use for each detection in the time axis. Detections
are plotted as "presence within some time bin" rather than attempting to plot the actual 
duration of a detected signal. 

```{r}
freqMap <- data.frame(
    type=paste0('Click_Detector_', 1:6),
    freqMin= c(40, 120, 1e3, 10e3, 3e3, 10),
    freqMax=c(200, 600, 2e3, 30e3, 6e3, 1e3)
)
plotAcousticScene(clicks, typeCol='detectorName', freqMap=freqMap, bin='30min')
```

Adjusting the timescale of the `bin` value can have a large affect on the appearance
of the plot. 

```{r}

plotAcousticScene(clicks, typeCol='detectorName', freqMap=freqMap, bin='1hour') /
        plotAcousticScene(clicks, typeCol='detectorName', freqMap=freqMap, bin='15min')
    
```

Similar to other frequency plots, `plotAcousticScene` also has a `scale` 
parameter to change whether frequency is shown in log or linear scale. Additionally,
the colors used can be manually specified by adding a `color` column to the `freqMap` input.

```{r}
freqMapManualColor <- freqMap
freqMapManualColor$color <- scales::viridis_pal()(nrow(freqMap))
plotAcousticScene(clicks, typeCol='detectorName', freqMap=freqMapManualColor, bin='15min', title='Log Scale') /
         plotAcousticScene(clicks, typeCol='detectorName', scale='linear', 
                           freqMap=freqMapManualColor, bin='15min', title='Linear Scale')
```

There is also an `alpha` parameter that allows users to set a transparency level for the plotted
regions, which can be useful for situations with many overlapping frequency and time ranges.

```{r}
plotAcousticScene(clicks, typeCol='detectorName', freqMap=freqMap, bin='15min', alpha=.8) /
    plotAcousticScene(clicks, typeCol='detectorName', freqMap=freqMap, bin='15min', alpha=.5)
```
