---
title: "Soundcoop2024 Examples"
author: "Taiki Sakai"
date: "2024-09-25"
output: html_document
---

# Install PAMscapes

Version 0.7.0 is available on CRAN, but there is also a 0.7.1 update from yesterday available 
on GitHub that has some more features. We'll also use the `here` package.

```{r}
install.packages('here')
install.packages('pak')
pak::pkg_install('TaikiSan21/PAMscapes')
# if above doesn't work, then try this next
# devtools::install_github('TaikiSan21/PAMscapes')
# if neither of those work, settle for CRAN version
# install.packages('PAMscapes')
```

```{r}
library(here)
library(PAMscapes)
```

The `here` package helps us deal with file paths relative to this RMarkown document. If
this file is in the same folder as your data, then `here('data1.nc')` will give the 
correct filepath to the file `'data1.nc'`. 

# Load Some Data

The function that handles all the data loading (regardless of format!) is
`checkSoundscapeInput`. You can give this a single file or multiple files
and it will load them up.

```{r}
# My data are in a folder NRS11 within another folder called testData
nrs11Folder <- here('testData/NRS11')
nrsFiles <- list.files(nrs11Folder, pattern='nc$', full.names=TRUE)
# I'll start with just the first 3 files
nrsData <- checkSoundscapeInput(nrsFiles[1:3])
```

If you updated to 0.7.1 from GitHub, this will automatically pull Lat/Long from the
.nc metadata, as well as pulling out the "platform" field. If you weren't able to,
lets add the Lat/Long so we can run some other examples

```{r}
nrsData$Longitude <- -123.45
nrsData$Latitude <- 37.88
```

# Download Wind & Rain Data

With Lat/Long added, we can match data from the Global Forecast System (GFS) weather
model, specifically wind and precipitation rate using the `matchGFS` funciton.

```{r}
nrsData <- matchGFS(nrsData)
```

This gives us new columns `windU`, `windV`, and `precRate`, lets convert
the directional winds to a single wind magnitude `windMag` since that is 
more relevant (I'll add this as a default output in the future). This
has also added columns `matchLong_mean`, `matchLat_mean`, and `matchTime_mean`
that just tell you waht coordinates and times in the environmental data
were matched with your data (it tries to match by closest available, these
columns are meant for troubleshooting / verifying).

```{r}
nrsData$windMag <- sqrt(nrsData$windU^2 + nrsData$windV^2)
```

Note that if you have any other environmental data you are interested in that
is available on an ERDDAP server, you can easily match that to your data
with the function `matchEnvData` in the `PAMmisc` package. Lets also grab
wave height from the WaveWatchIII wave model available on ERDDAP. The variable
`Thgt` is the significant wave height, it will get stored in our data as
`Thgt_mean`.

```{r}
library(PAMmisc)
nrsData <- matchEnvData(nrsData, nc='ww3_global', var=c('Thgt'))
```

Lets also create some general categories for our wind data that we can use to explore 
the data. We'll create categories `0-10`, `10-20`, and `20+` (arbitrarily).

```{r}
nrsData$windCategory <- cut(nrsData$windMag, breaks=c(0, 10, 20, 100), 
                            labels=c('Low Wind', 'Medium Wind', 'High Wind'))
```


# Plot Exploration

Okay enough of that, lets look at some plots. New as of version 0.7.0
there's a `shiny` app that lets you easily explore some example plots
with your data, `runSoundscapeExplorer`. So we just have to run that,
giving it the data we want to explore.

```{r}
runSoundscapeExplorer(nrsData)
```

# AIS Data 

This part will download some *very* large files and may take quite a while, so run at your own risk.
The `downloadMarCadAIS` function will download a 250mb .zip file and 700mb for *every single day*
in your dataset. Marine Cadastre stores their data as one file containing all geographic regions
for a single day. So if your dataset covers many days, this download will be a sad time. Giving
this function your data it will sort out which files to download, and store them in `outDir`.

```{r}
marcadFiles <- downloadMarCadAIS(nrsData, outDir = here('AIS'), overwrite = FALSE, unzip=TRUE)
```

These files are extremely unwieldy to work with, and you probably don't need data covering the
entire geographic region for your project. So the `subsetMarCadAIS` funciton will take those
full size 700mb files and create a smaller working dataset. In this case lets pretend that all
the recorders I work with are on the west coast, so I'll create one working AIS directory that 
contains AIS data for only the west coast data. This reduces each daily file down to ~130mb instead
of 700mb, which is still large but much easier to handle.

```{r}
subsetMarCadAIS(inDir=here('AIS'), outDir=here('AIS_West'), name='West_', 
                latRange=c(20,50), lonRange=c(-140,-110),
                overwrite=FALSE, progress=TRUE)
```

Now we can match some of this AIS data to our data. We just need to define what counts as a
"close" vessel event for our recorder, and we can get some summary information about vessels
that pass within a certain distance of our recorders. The `distance` argument here is in
meters, so here we'll get information about vessels passing within 10km of our recorder.
`ais` is the folder path of our smaller "working" AIS dataset.

This function adds columns:
- `nShips` the number of vessels within `distance` at this time
- `meanDist` the average distance (meters) to nearby ships (`NA` if `nShips` is 0)
- `meanSOG` the average speed (knots) over ground of the nearby ships (`NA` if `nShips` is 0)
- `closeDist` the distance (meters) to the closest nearby ship (`NA` if `nShips` is 0)
- `closeSOG` the speed over ground (knots) of the closest nearby ship (`NA` if `nShips` is 0)

```{r}
nrsData <- addAISSummary(nrsData, ais=here('AIS_West'), distance=10e3)
```
