---
title: "AIS Workflow Example"
author: "Taiki Sakai"
output: html_document
---

## Introduction

This document is a brief outline to using the new AIS download / matching functions in the
`PAMscapes` R package. All functions use AIS data downloaded from [MarineCadastre](https://marinecadastre.gov/data/).
Data from other sources may work with some of the functions, but formatting may need to be adjusted.

Currently this package is only available on GitHub, and can be installed with:

```{r, eval=FALSE}
devtools::install_github('TaikiSan21/PAMscapes')
```

Load packages for the start of the examples:

```{r}
library(here)
library(PAMscapes)
library(readr)
```
### Download from Marine Cadastre

For this example we will use drifter track data from one of the SWFSC ADRIFT buoy deployments. The first step
is to read in the data you want to download AIS data for. The requirements are that this data have columns
"UTC", "Longitude", and "Latitude", and that the "UTC" column is converted to `POSIXct`. Here we'll use
`readr::read_csv` to handle the type conversion for us. This drift also has a couple wonky spots at the start,
so we are dropping the first two rows.

```{r, cache=TRUE}
gps <- read_csv(here('devel/ADRIFT_006_GPS.csv'))
gps <- gps[-(1:2),]
```

The function `downloadMarCadAIS` will look at your data, and start downloading the appropriate AIS
files from Marine Cadastre (MC). The files on MC are stored as daily files covering the entire geographic
range, so are very large and cumbersome to deal with. They download as .zip files which contain a single large
(700mb +) CSV file. `downloadMarCadAIS` will download the appropriate files covering your dataset, and unzip them 
to CSV if `unzip=TRUE`. An important note is that if you use the same `outDir` folder for all of your AIS storage,
the function will check if the file exists first before trying to download a new file from MC if `overwrite=FALSE`.
This is the default option and should almost always be used, unless something has gone wrong with a previous attempt.
This function can take quite a while depending on the number of days you need to download, it will show a progress
bar so that you know its working.

```{r, cache=TRUE}
marcadFiles <- downloadMarCadAIS(gps, outDir = here('AIS'), overwrite = FALSE, unzip=TRUE)
```

The full size files are not needed for any specific analysis, so it is often useful to subset these files to a
geographic region that you work with often. The function `subsetMarCadAIS` lets you create a geographic subset
of the AIS data that is smaller and easier to work with. I work with US West Coast data, so I'll specify a very rough
Lat/Long boundary to create a new folder called "AIS_West". These will go through each file in `inDir` and subset it
to those boundaries, and then create a new file in `outDir` that has the same file name but prefixed with "West_"
(`name='West_'`). This function can also take quite a while, so it is recommended to run with `overwrite=FALSE` so
that if a file has already been subset in the past it will not read and subset it again.

```{r, cache=TRUE}
subsetMarCadAIS(inDir=here('AIS'), outDir=here('AIS_West'), name='West_', 
                latRange=c(20,50), lonRange=c(-140,-110),
                overwrite=FALSE, progress=TRUE)
```

### Read AIS Data Into R

Now we have a folder "AIS_West" that contains smaller (~140mb vs ~700mb) files to work with that are relevant
to our study region. This is still too much data to really work with, and most of it is not relevant to our
specific GPS for this example. Instead of simply reading in all of these, we can use the `readLocalAIS` function
to read in only AIS data that come within a certain distance of our buoys. Here we set `distance=10e3` to read in
only AIS data for ships that come within 10km of our drifter track. This function will read in only AIS data that
fall within the time range of the "UTC" values in `gps`, if additional data before/after that are desired you 
can specify the `timeBuff` argument (units seconds) to capture additional time before and after the time range
of your GPS data. This can be useful to show the full picture of ship activity just before/after your data.

```{r, cache=TRUE}
ais <- readLocalAIS(gps, aisDir=here('AIS_West'), distance=10e3)
str(ais)
```

Here's a list of the columns in `ais`:

- `Latitude` - Latitude coordinate of the ship
- `Longitude` - Longitude coordinate fo the ship
- `MMSI` - MMSI Id of the ship
-  `UTC` - Timestamp of coordinate update
- `vesselLength` - Length of ship (I believe this is meters)
- `SOG`  - Speed Over Ground of ship
- `buoyLat` - Latitude coordinate of the input `gps` dataframe at time `UTC`
- `buoyLong` - Longitude coordinate of the input `gps` dataframe at time `UTC`
- `buoyDist` - Distance (meters) between the Latitude/Longitude coordinate and buoyLat/buoyLong coordinates
- `inDist` - Logical flag whether or not `buoyDist` is less than the input `distance` (10e3 in this case)
- `group` - Individual vessel tracks are marked as different groups when the vessel travels in/out of range.
This is mostly used internally by some testing functions while I was developing the code.

Now you have AIS data! Note that this contains a lot of AIS data that will not be particularly close
to your input GPS data. It contains **all** AIS data from the entire time range of `gps` for any
vessel that traveled within `distance=10e3` of our buoy for even a short amount of time. Since ships
move fast and far, a lot of this is quite far from our buoy. The plot below shows the full extent of
AIS data we have loaded.

```{r}
library(ggplot2)
ggplot() +
    geom_path(data=gps, aes(x=Longitude, y=Latitude), col='black') +
    geom_path(data=ais, aes(x=Longitude, y=Latitude, group=factor(MMSI), col=factor(MMSI)))
```

A more relevant plot would be to truncate this to only data within the Lat/Long bounds of our GPS data.
We can also color the tracks by whether or not they were within 10km of our buoy at that time (bottom 
plot).

```{r}
library(patchwork)
(ggplot() +
    geom_path(data=gps, aes(x=Longitude, y=Latitude), col='black') +
    geom_path(data=ais, aes(x=Longitude, y=Latitude, group=factor(MMSI), col=factor(MMSI))) +
    xlim(min(gps$Longitude), max(gps$Longitude)) +
    ylim(min(gps$Latitude), max(gps$Latitude))
) /
    (ggplot() +
    geom_path(data=gps, aes(x=Longitude, y=Latitude), col='black') +
    geom_path(data=ais, aes(x=Longitude, y=Latitude, group=factor(MMSI), col=inDist)) +
    xlim(min(gps$Longitude), max(gps$Longitude)) +
    ylim(min(gps$Latitude), max(gps$Latitude)))
```

### Match AIS to Our Data

`ais` contains a lot of data we don't need, but we might be more interested in connecting this
data specifically back to our original `gps` dataframe. The function `addAIS` will do this,
pairing the times in `gps` with the AIS data for each vessel in `ais`. However, this gets a bit
tricky because of the differing timescales of `gps` and `ais`. Our buoy coordinate updates are
often quite far apart (30-60 minutes) relative to the speed of the AIS updates. In the time between
GPS updates most vessels will travel quite far, so just connecting our GPS times to AIS data
as-is can be a bit confusing. The solution to get around this is to use interpolation to make
our GPS data have a similar timescale to the AIS data we are matching to. `addAIS` allows for three
different types - "none" does no interpolation, "close" will only interpolate GPS coordinates at times
when the vessel is marked as `inDist=TRUE` from earlier, and "all" will simply interpolate the full
time range of our GPS data. The parameter `interpTime` controls how far apart to space the new coordinates.

The main reason one might want to use `interpType='close'` instead of `interpType='all'` is that the 
`addAIS` function multiplies the number of rows in your data by quite a bit. Each separate vessel in `ais`
will be paired with each coordinate in `gps`, so if there are 5 unique vessels in `ais` then the result
of `addAIS` will contain 5x as many rows as `gps`. Interpolating the coordinate points further multiplies
the number of rows in the output, so it can get quite large. You can see below that we start with only
222 rows of points, but `interpType='all'` and `interpTime=120` ends up with 72,838 rows. This is probably
fine for most datasets, but may be prohibitive for larger datasets.

```{r}
gpsNone <- addAIS(gps, ais, interpType = 'none')
gpsClose <- addAIS(gps, ais, interpType = 'close', interpTime = 120)
gpsAll <- addAIS(gps, ais, interpType = 'all', interpTime = 120)
cat('Original:', nrow(gps),
    'NoInterp:', nrow(gpsNone),
    'CloseInterp:', nrow(gpsClose),
    'AllInterp:', nrow(gpsAll))
```
The plots below illustrate the difference in output - no interpolate results in quite a 
confusing picture of ship activity due to the different timescales. "close" interpolation
gives an accurate image of activity, but only when ships are nearby. "all" interpolation
gives a more complete picture of ship activity as they are passing by our buoy. 

```{r}
baseGps <- ggplot(data=gps, aes(x=Longitude, y=Latitude), col='black') +
    geom_path() +
    xlim(min(gps$Longitude), max(gps$Longitude)) +
    ylim(min(gps$Latitude), max(gps$Latitude))
(baseGps +
    geom_path(data=gpsNone, aes(x=shipLong, y=shipLat, group=factor(MMSI), color=shipDist < 10e3)) +
        geom_point(data=gpsNone, aes(x=shipLong, y=shipLat, color=shipDist < 10e3)) +
    ggtitle('NoInterp')) /
(baseGps +
    geom_path(data=gpsClose, aes(x=shipLong, y=shipLat, group=factor(MMSI), color=shipDist < 10e3)) +
     geom_point(data=gpsClose, aes(x=shipLong, y=shipLat, color=shipDist < 10e3)) +
    ggtitle('CloseInterp')) /
(baseGps +
    geom_path(data=gpsAll, aes(x=shipLong, y=shipLat, group=factor(MMSI), color=shipDist < 10e3)) +
     geom_point(data=gpsAll, aes(x=shipLong, y=shipLat, color=shipDist < 10e3)) +
    ggtitle('AllInterp'))
```

There are some other important things to consider when using interpolation with `addAIS`. It
is necessary that some columns of the input `gps` will be lost. Comparing the column names
of the input and output:

```{r}
print(colnames(gps))
cat('\n')
print(colnames(gpsAll))
```

Some columns like "DriftName" and "DeploymentSite" were kept in the output, but others like "recordingEffort"
and "knots" were not. Any columns in the input `gps` that contain only a single unique value will be kept in the 
output, but all other columns will be dropped. This is because `addAIS` does not know how to handle other data types
when interpolating. If there are some columns present that would be appropriate to linearly interpolate between GPS
points, then these column names can be given to the `interpCols` parameter, and they will then be present in the output.
As an example, we'll use `matchEnvData` from the `PAMmisc` package to download some sea-surface temperature data for
our GPS input.

```{r, cache=TRUE}
library(PAMmisc)
gps <- matchEnvData(gps, nc='jplMURSST41', var='analysed_sst')
str(gps)
```

We now have a column "analysed_sst_mean" that should be appropriate to interpolate, so we can give this
as the `interpCols` argument and it will be kept in the output:

```{r}
gpsAllSST <- addAIS(gps, ais=ais, interpType='all', interpTime=120, interpCols='analysed_sst_mean')
str(gpsAllSST)
```
