Load some data and make a fake variable for "by" plotting
```{r}
library(PAMscapes)
library(here)
data <- loadSoundscapeData(here('testData/NRS11/NRS11_20200101.nc'))
data$hour <- lubridate::hour(data$UTC)
data$night <- data$hour > 16
```
Original version looked like this

```{r}
plotPSD(data, q=.05, by='night')
```

Can adjust the line widths for both the median line and the quantile borders with 
`lwd` and `borderLwd`, respectively. Defaults are 1 and 0.5

```{r}
plotPSD(data, q=.05, by='night', lwd=2, borderLwd=1)
```

Can change the linetype used, can either be the numeric 0-6 code or name 
(see [https://ggplot2.tidyverse.org/reference/aes_linetype_size_shape.html]).
If a single value is provided it will be used for all levels, otherwise yo must
provide as many linetypes as there are levels.

```{r}
plotPSD(data, by='night', q=.05, linetype=3, title='All same linetype')
# these specify the same linetype mapping
plotPSD(data, by='night', q=.05, linetype=c(1, 2), title='Different types (by number)')
plotPSD(data, by='night', q=.05, linetype=c('solid', 'dashed'), title='Different types (by name)')
plotPSD(data, by='night', q=.05, linetype=c(0, 1), title='linetype 0/blank makes lines disappear')
```

To make it easier to assign specific colors and linetypes to specific variable values,
both `color` and `linetype` can now accept a named vector where the name is the variable
value. Previously you had to just assign them in the correct order.

```{r}
plotPSD(data, by='night', q=.05, color=c('red', 'blue'), title='Color by order')
plotPSD(data, by='night', q=.05, color=c('TRUE'='red', 'FALSE'='blue'), title='Color by name')
plotPSD(data, by='night', q=.05, linetype=c(1, 3), title='Linetype by order')
plotPSD(data, by='night', q=.05, linetype=c('TRUE'=1, 'FALSE'=3), title='Linetype by name')
```

I also added an option to turn off the "nObs (###)" from the legend if you dont want
it there with `showNObs=FALSE`

```{r}
plotPSD(data, by='night', showNObs=FALSE, title='No "nObs"')
```

Added an option to change the opacity of the color fill when plotting a quantile 
with `alpha`, the default value is 0.1
```{r}
plotPSD(data, by='night', q=0.05, alpha=.5, title='More opaque shading')
```

