---
title: 'Soundscape Plot Examples'
output: html_document
---
# Soundscape Plots

Load raw soundscape data

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(here)
tolFile <- here('testData/ADRIFT_017_TOL_mean_2min.csv')
olFile <- here('testData/ADRIFT_017_OL_mean_2min.csv')
bbFile <- here('testData/ADRIFT_017_BB_mean_2min.csv')
```
Create function to read triton data and format into long-style to be useful

```{r}
library(tidyr)
readTriton <- function(x) {
    data <- read_csv(x, show_col_types=FALSE)
    if(!colnames(data)[1] == "yyyy-mm-ddTHH:MM:SSZ") {
        warning('Format does not appear to be from Triton, formatting stopped')
        return(data)
    }
    colnames(data)[1] <- 'UTC'
    data
}

tritonToLong <- function(x) {
    type <- gsub('_[0-9-]+', '', colnames(x)[2])
    if(type != 'BB') {
        x <- pivot_longer(x, cols=2:ncol(x), names_to='type', values_to='value')
        x$frequency <- as.numeric(gsub('[A-z]+_', '', x$type))
        x$type <- gsub('_[0-9-]+', '', x$type)
    }
    if(type == 'BB') {
        freqs <- as.numeric(strsplit(gsub('BB_', '', colnames(x)[2]), '-')[[1]])
                 # as.numeric(strsplit(gsub('BB_', '', colnames(x)[2]), '-')[[1]])
        x$type <- 'BB'
        colnames(x)[grepl('BB_', colnames(x))] <- 'value'
        x$freqMin <- freqs[1]
        x$freqMax <- freqs[2]
    }
    x
}
ol <- readTriton(olFile)
tol <- readTriton(tolFile)
bb <- readTriton(bbFile)
olLong <- tritonToLong(ol)
tolLong <- tritonToLong(tol)
bbLong <- tritonToLong(bb)
```
## Slide 3c

Frequency (x) vs Hour of day (y) using TOL or OL as input
Units for PSD are "dB re: 1 uPa^2/Hz"
Units for OL, TOL, and BB are "dB re: 1uPa"
```{r}
library(lubridate)
# needs UTC, frequency, value
slide3c <- function(x, title=NULL, scale=c('log', 'linear'), freqMin=NULL, toTz='UTC',
                    cmap=scales::viridis_pal()(25)) {
    scale <- switch(match.arg(scale),
                    'log' = 'log10',
                    'identity'
    )
    x <- tritonToLong(x)
    # x <- tolLong
    x$UTC <- with_tz(x$UTC, tzone=toTz)
    x$hour <- hour(x$UTC)
    summByHour <- group_by(x, hour, frequency) %>% 
        summarise(value = median(value), .groups='drop')
    summByHour$hour_end <- summByHour$hour + 1 # hour ranges 0-23
    freqVals <- sort(unique(summByHour$frequency))
    freqDiffs <- diff(freqVals)
    freqDiffs <- c(freqDiffs[1], freqDiffs)
    names(freqDiffs) <- as.character(freqVals)
    summByHour$freq_low <- summByHour$frequency - freqDiffs[as.character(summByHour$frequency)]
    summByHour$freq_low <- ifelse(summByHour$freq_low < .0001, .0001, summByHour$freq_low)
    if(is.null(title)) {
        title <- x$type[1]
    }
    if(is.null(freqMin)) {
        freqMin <- min(summByHour$freq_low)
    }
    if(freqMin < 1 && scale == 'log10') {
        freqMin <- 1
    }
    ggplot(summByHour) +
        geom_rect(aes(ymin=hour, 
                      ymax=hour_end,
                      xmin=freq_low,
                      xmax=frequency,
                      fill=value)) +
        scale_x_continuous(trans=scale, expand=c(0,0), limits=c(freqMin, max(freqVals))) + 
        scale_y_continuous(expand = c(0,0)) +
        scale_fill_gradientn(colors=cmap) +
        ggtitle(title) +
        labs(x='Frequency (Hz)', 
             y=paste0('Hour (', toTz, ')'),
             fill = 'dB re: 1uPa') +
        theme(legend.title = element_text(angle=90)) +
        guides(fill=guide_colorbar(title.position='right', barheight=10, title.hjust=.5))
}
slide3c(tol, toTz='UTC')
```

```{r}
library(patchwork)
slide3c(tol, title='TOL Log scale', freqMin=10) / slide3c(tol, title='TOL Linear scale', freqMin=10, scale='linear')
# slide3c(tol, title='TOL linear scale', scale='linear') / slide3c(ol, scale='linear')
```
Time series plots - we want confidence interval?? we have every 2 minutes, so guess we need
to bin it to some bigger unit then plot time series
```{r}
# slim is placeholder to make minimal embelishment plot to fit into other plots
# needs UTC, then user defined column and by
slide6 <- function(x, bin='1hour', column, ci=0, title=NULL, by=NULL, slim=FALSE) {
    x$timeBin <- floor_date(x$UTC, unit=bin)
    if(length(ci) == 1) {
        ci <- c(ci, ifelse(ci==0, 0, 1-ci))
    }
    ci <- sort(ci)
    if(is.null(by)) {
        plotData <- group_by(x, timeBin)
    } else {
        x[[by]] <- factor(x[[by]])
        plotData <- group_by(x, timeBin, .data[[by]])
    }
    if(!all(ci == 0)) {
        plotData <- plotData %>% 
            summarise(med = median(.data[[column]]),
                      qlow = quantile(.data[[column]], probs=ci[1]),
                      qhigh = quantile(.data[[column]], probs=ci[2]),
                      .groups='drop')
    } else {
        plotData <- plotData %>% 
            summarise(med = median(.data[[column]]),
                      .groups='drop')
    }
    if(is.null(by)) {
        g <- ggplot(data=plotData, aes(x=timeBin))
    } else {
        g <- ggplot(data=plotData, aes(x=timeBin, col=.data[[by]], fill=.data[[by]]))
    }
    g <- g + geom_line(aes(y=med))
    if(!all(ci == 0)) {
        g <- g + 
            geom_ribbon(aes(ymin=qlow, ymax=qhigh), alpha=.1)
    }
    if(is.null(title)) {
        title <- column
    }
    g <- g + ggtitle(title) +
        labs(x='Date', y='SPL (dB re: 1uPa)')
    g
}
```

```{r}
gps <- read_csv(here('testData/ADRIFT_017_GPS.csv'))
slide6(tol, bin='2hour', column='TOL_160') /
    slide6(tol, bin='2hour', column='TOL_160', ci=.05, title='TOL160 with 95% quantiles')
(ggplot(gps, aes(x=Longitude, y=Latitude)) +
    geom_path()) /
    slide6(tol, bin='2hour', column='TOL_160') +
    plot_layout(heights=c(4,1))

```
```{r}
# needs UTC, thne user defined column
slide5 <- function(x, column, title=NULL, cmap=scales::viridis_pal()(25)) {
    x$hour <- hour(x$UTC)
    x$day <- floor_date(x$UTC, unit='1day')
    if(is.null(title)) {
        title <- column
    }
    plotData <- group_by(x, day, hour) %>% 
        summarise(med = median(.data[[column]]), .groups='drop')
    ggplot(plotData) +
        geom_rect(aes(xmin=hour, xmax = hour + 1,
                         ymin=day, ymax = day + 3600*24,
                      fill=med)) +
        scale_fill_gradientn(colors=cmap) +
        scale_x_continuous(limits=c(0,24), expand=c(0,0)) +
        labs(x='Hour', 
             y='Date',
             fill = 'dB re: 1uPa') +
        theme(legend.title = element_text(angle=90)) +
        guides(fill=guide_colorbar(title.position='right', barheight=10, title.hjust=.5)) +
        ggtitle(title)
}
slide5(tol, column='TOL_160')
```

Acoustic scene boxy plot needs freq ranges
```{r, cache=TRUE}
freqMap <- data.frame(
    type=paste0('Click_Detector_', 1:6),
    freqMin= c(40, 120, 1e3, 10e3, 3e3, 10),
    freqMax=c(200, 600, 2e3, 30e3, 6e3, 1e3),
    color=c('red', 'blue', 'green', 'darkgreen', 'black', 'black')
)
data <- readRDS(here('testData/adrift17study.rds'))
library(PAMpal)
clicks <- getClickData(data)
clicks$species <- clicks$detectorName
```

```{r}
# helper to make periods out of bin units
unitToPeriod <- function(x) {
    x <- gsub('([0-9]*)(.*)', '\\1_\\2', x)
    x <- strsplit(x, '_')[[1]]
    if(x[1] == '') {
        x[1] <- '1'
    }
    period(as.numeric(x[1]), units=x[2])
}
# needs UTC, species==type, 
slide7 <- function(x, freqMap, title=NULL, bin='1day', scale=c('log', 'linear'), freqMin=NULL) {
    x$plotStart <- floor_date(x$UTC, unit=bin)
    thisPeriod <- unitToPeriod(bin)
    x$plotEnd <- x$plotStart + thisPeriod
    scale <- switch(match.arg(scale),
                    'log' = 'log10',
                    'identity'
    )
    x <- left_join(x, freqMap, by=join_by('species'=='type'))
    x <- x[!is.na(x[['freqMin']]), ]
    x <- distinct(x[c('plotStart', 'plotEnd', 'freqMin', 'freqMax', 'species')])
    if(is.null(freqMin)) {
        freqMin <- min(x[['freqMin']])
    }
    if(freqMin < 1 && scale == 'log10') {
        freqMin <- 1
    }
    g <- ggplot(x) +
        geom_rect(aes(xmin=plotStart,
                      xmax=plotEnd,
                      ymin=freqMin,
                      ymax=freqMax,
                      fill=species)) +
        scale_y_continuous(trans=scale) +
        labs(y='Frequency (Hz)',
             x='Date',
             fill='Sound Type')
    if('color' %in% colnames(freqMap)) {
        colNames <- freqMap[['color']]
        names(colNames) <- freqMap[['type']]
        g <- g +
            scale_fill_manual(values=colNames)
    }
    g
}
s7 <- slide7(clicks, bin='15min', freqMap=freqMap, scale='log')
s7
```
